Allometric growth parameters for invasive lionfish \emph{Pterois volitans} (Actinopterygii, Scorpaenidae) in the Western Atlantic

### 2018-02-24 Manuscript.tex
Need to work a lot more on this, the reviewers gave good comments but it needs A LOT more work

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{\Large{Weight-at-length of the invasive lionfish \textit{Pterois volitans} (Actinopterygii, Scorpaenidae) in the Central Mexican Caribbean, and a review of allometric growth parameters across the invasion range}}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
\subtitle{\textbf{Running title}: Length-Weight relationships of lionfish}
  \author{}
  \preauthor{}\postauthor{}
  \date{}
  \predate{}\postdate{}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage[table]{xcolor}
\usepackage{wrapfig}
\usepackage{setspace}
\doublespacing

\begin{document}
\maketitle

Juan Carlos Villaseñor-Derbez\textsuperscript{1}

\textsuperscript{1} Bren School of Environmental Sciences and
Management, University of California Santa Barbara, Santa Barbara,
California, U.S.

* Correspondence: Juan Carlos Villaseñor-Derbez, Bren Hall, University
of California at Santa Barbara, Santa Barbara, CA, 93106, US phone: +1
207 205 8435, e-mail:
\href{mailto:jvillasenor@bren.ucsb.edu}{\nolinkurl{jvillasenor@bren.ucsb.edu}}

\textbf{Abstract}

Lionfish (\emph{Pterois volitans/miles}) are an invasive species in the
North-Western Atlantic and the Caribbean. In order to better manage the
invasion, inform lionfish removal programs, and estimate biomass
available for harvest, we must be able to accurately estimate their
total biomass, frequently from length observations. This work compares
length-weight relationships of the invasive lionfish through the
invasion range and reports the length--weight relationship for lionfish
in the Central Mexican Caribbean. A review of 13 length--weight
relationships reported in eight peer-reviewed studies and FishBase is
provided. These parameters were used to identify spatial variation in
weight-at-length. For a given length, parameters from the Caribbean
yielded lower weights than those from the Gulf of Mexico and Atlantic,
indicating that weight-at-length is spatially variable. This highlights
the importance of using site-specific parameters to estimate biomass
from length observations. This study also reports a new pair of
length-weight parameters (\(a = 3.2056\times 10^{-6}; b = 3.235\)) for
organisms sampled in the Central Mexican Caribbean. Findings from this
work can aid managers and decision makers to better select length-weight
parameters when these are not available for their region of interest.

\textbf{Resumen}

El pez león (\emph{Pterois volitans/miles}) es una especie invasora en
el Atlántico Noroeste y el Mar Caribe. Para tener un mejor manejo de la
invasión, informar programas de remoción y estimar la biomasa disponible
para aprovechamiento, debemos ser capaces de estimar la biomasa
disponible con presisión, frecuentemente a partir de observaciones de
tallas. Este trabajo compara la relación longitud--peso del pez león a
través del rango de invasión, y reporta un nuevo par de parámetros de
crecimiento alométrico para la zona central del Caribe Mexicano. El
trabajo presenta revisión de 13 relaciones longitud--peso reportadas
reportadas en ocho estudios publicados en revistas arbitradas y
FishBase. Para una misma talla, los parámetros del Caribe indican un
mejor peso que los parámetros del Golfo de México y el Atlántico,
indicando una variabilidad espacial en la relación longitud peso. Esta
variabilidad resalta la importancia de utilizar parámetros específicos a
un sitio al estimar biomasa a partir de observaciones de tallas. Este
estudio también reporta una nueva relación longitud--peso
(\(a = 3.2056\times 10^{-6}; b = 3.235\)) para organismos muestreados en
la costa central del Caribe Mexicano. Los resultados de este trabajo
pueden ayudar a tomadores de decisiones en la selección de relaciones
longitud-peso cuando éstas no están disponibles para su zona de interés.

\textbf{Key words}\\
biometry, invasive species, spatial variability, allometric growth,
Mexico

@@

At least 84\% of the marine eco-regions have reported the presence of an
invasive species (Molnar et al., 2008). These represent a major threat
to local biodiversity and the economic activities that depend on it,
like tourism or fisheries (Bax et al., 2003). Invasive species may also
threaten native species through competition (DAVIS, 2003) or predation.
By 2005, the economic cost of invasive species to the United States was
estimated at \$120 billion per year and nearly 42\% of species that have
been included in the Endangered or Threatened species list have been
labeled as such due to presence of invasive species (Pimentel, Zuniga \&
Morrison, 2005). This highlights the importance of understanding,
managing, and preventing ecological invasions.

Lionfish (\emph{Pterois volitans/miles} complex) are an invasive species
in the North-Western Atlantic and the Caribbean, likely introduced
through liberation of aquarium-kept organisms (Betancur-R et al., 2011).
They are the first marine vertebrates to establish in North Atlantic
(Schofield, 2009, 2010) and Caribbean coasts (Sabido-Itza et al., 2016).
Lionfish have been widely reported in coral reefs (Aguilar-Perera \&
Tuz-Sulub, 2010), but also in other habitats such as estuaries (Jud et
al., 2011), mangroves (Barbour et al., 2010), areas with hard-bottoms
(Muñoz, Currin \& Whitfield, 2011), and mesophotic reefs (Andradi-Brown
et al., 2017). Due to its threat to local biodiversity, the speed of
their spread, and its difficulty of management, their presence in these
waters has been labeled as a major marine invasion (Hixon et al., 2016).

A significant amount of research has been done to describe lionfish
feeding ecology in North Carolina (Muñoz, Currin \& Whitfield, 2011),
the Bahamas (Morris \& Akins, 2009; Cote et al., 2013), Northern Gulf of
Mexico (Dahl \& Patterson, 2014), Mexican Caribbean (Valdez-Moreno et
al., 2012; Villaseñor-Derbez \& Herrera-P\a'erez, 2014), Belize
(Hackerott et al., 2017), and Costa Rica (Sandel et al., 2015). Their
feeding behavior and high consumption rates can reduce recruitment
(Albins \& Hixon, 2008) and population sizes (Green et al., 2012) of
native reef-fish species, and further the endangerment of critically
endangered reef fish (Rocha et al., 2015). (However, see Hackerott et
al. (2017) for a case where there was no evidence that lionfish affected
the density, richness, or composition of prey fishes). Major efforts
have also been made to understand the possible impacts of the invasion
by keeping track of its range through time (Schofield, 2009, 2010) and
predicting invasion ranges under climate change scenarios (Grieve,
Curchitser \& Rykaczewski, 2016). By combining information from these
disciplines, researchers have been able to predict the trophic impacts
of lionfish (Arias-Gonzalez et al., 2011), which can then be translated
into ecosystem-level and economic impacts.

Seeking to reduce lionfish densities, governments and non-profit
organizations have promoted removal programs and incentivized its
consumption (Chin, Aiken \& Buddo, 2016). In some cases, these have
shown to significantly reduce -but not quite eliminate- lionfish
abundances at local scales (Sandel et al., 2015, Chin, Aiken \& Buddo
(2016), de Leon et al. (2013)). The rapid recovery rates exhibited by
lionfish (Barbour et al., 2011) and the persistent populations in
mesophotic coral ecosystems (Andradi-Brown et al., 2017) -which can
contribute with recruitment to shallow-water populations- make of
complete eradication through fishing effort an unlikely solution.
However, further incentivizing its consumption might create a demand big
enough to promote and sustain a stable fishery (Chin, Aiken \& Buddo,
2016), which can reduce local abundances and control -not eradicate- the
invasion while providing alternative livelihoods.

The feasibility of lionfish removal programs has been extensively
evaluated through field observations(Sandel et al., 2015, Chin, Aiken \&
Buddo (2016), de Leon et al. (2013); Usseglio et al., 2017) and
empirical modeling (Barbour et al., 2011; Morris, Shertzer \& Rice,
2011; Johnston \& Purkis, 2015). The latter measure changes in biomass
or density (Barbour et al., 2011; Johnston \& Purkis, 2015) in response
to increased mortality (\emph{i.e.} lionfish removal). In this case,
biomass represents the sum of all fish's individual weight. Total Weight
(TW) can be estimated from Total Length (TL) observations using the
allometric growth equation (\(TW = aTL^b\)). Parameters \(a\) and \(b\)
for this equation exist for North Carolina (Barbour et al., 2011),
Northern (Fogg et al., 2013) and Southern Gulf of Mexico (Aguilar-Perera
\& Quijano-Puerto, 2016), the Southern Mexican Caribbean (Sabido-Itza et
al., 2016), Little Cayman (Edwards, Frazer \& Jacoby, 2014), Jamaica
(Chin, Aiken \& Buddo, 2016), Bonaire (de Leon et al., 2013) and Costa
Rica (Sandel et al., 2015), but remain unavailable for the central
Mexican Caribbean. The weight-at-length of a species can vary across
regions as a response to biotic (\emph{e.g.} local food availability)
and abiotic (\emph{e.g.} water temperature) conditions (Johnson \&
Swenarton, 2016). Thus, when using biomass-informed models or estimating
biomass from length observations, it is important to use site-specific
parameters to obtain an accurate estimate. This is especially important
when research involves identifying the total biomass available for
harvest by fishers (Chin, Aiken \& Buddo, 2016) or the efficacy of
lionfish removals (Barbour et al., 2011; Morris, Shertzer \& Rice, 2011;
Johnston \& Purkis, 2015).

@@

The study took place off the coasts of Playa del Carmen, in the Mexican
Caribbean (\textbf{Fig. 1}). The region represents the northernmost
section of the Mesoamerican Barrier Reef System (Ruiz-Zarate \&
Arias-Gonzalez, 2004). Coral reefs and mangroves are locally important
habitats that represent important sources of income in terms of
extractive (\emph{e.g.} recreational fishing) and non-extractive
(\emph{e.g.} SCUBA diving) activities related to tourism, the main
source of income to the local economy (Murray, 2007).

The reef profile has been described by Arias-Gonzalez (1998), indicating
that the reef lagoon extends about 500 m from the coast, until the reef
crest is reached. The reef becomes deeper, leading to the reef front
often found at 700 m from the coastline and extends for an additional
300 m. At approximately 1000 m away from shore and 30 - 40 m depth, the
reef leads to a drop-off. Along a perpendicular profile to the coast,
bands of reef are interrupted by sand patches at 8 - 12 m deep and 16-18
m deep. Along the coast, these reefs have been reported to be under
significant anthropogenic pressure, likely causing a shift in structure
and function (Bozec et al., 2008).

\begin{figure}[htbp]
\centering
\includegraphics{Manuscript_files/figure-latex/create map-1.pdf}
\caption{Map of the study area. The black inset on the left (Yucatan
Peninsula) indicates the location where study sites are distributed. On
the right, circular markers indicate sampling sites and the black
romboid indicates location of Puerto Aventuras, Mexico.}
\end{figure}

@@

\begin{longtable}[]{@{}llllllr@{}}
\caption{Coordinates, minimum, maximum and mean depth (m), and number of
samples for each location.}\tabularnewline
\toprule
Location & Lat. & Long. & Min. Depth & Max. Depth & Mean Depth &
n\tabularnewline
\midrule
\endfirsthead
\toprule
Location & Lat. & Long. & Min. Depth & Max. Depth & Mean Depth &
n\tabularnewline
\midrule
\endhead
Canones & 20.477 & -87.233 & 15.0 & 31.2 & 21.6 & 11\tabularnewline
Castillo & 20.496 & -87.220 & 12.5 & 30.5 & 27.5 & 18\tabularnewline
Cuevitas & 20.478 & -87.244 & 7.4 & 12.8 & 11.2 & 4\tabularnewline
Islas & 20.490 & -87.228 & 14.0 & 19.4 & 16.7 & 10\tabularnewline
Paamul & 20.513 & -87.192 & 9.9 & 22.7 & 15.5 & 31\tabularnewline
Paraiso & 20.484 & -87.226 & 9.4 & 38.1 & 17.7 & 16\tabularnewline
Pared & 20.502 & -87.212 & 12.1 & 21.0 & 16.3 & 12\tabularnewline
Pedregal & 20.507 & -87.204 & 14.4 & 14.9 & 14.7 & 3\tabularnewline
Santos & 20.493 & -87.222 & 5.7 & 26.6 & 16.2 & 2\tabularnewline
Tzimin-Ha & 20.393 & -87.307 & 21.2 & 24.6 & 22.9 & 2\tabularnewline
Total & & & 5.7 & 38.1 & 18.6 & 109\tabularnewline
\bottomrule
\end{longtable}

@@

\[TW = aTL^b\]

@@

\[log_{10}(TW) = b\times log_{10}(TL) + log{10}(a)\]

@@

\[Y = mX + c\]

@@

Other length-weight relationships (n = 13) were extracted from
peer-reviewed literature. Parameters were obtained for North Carolina (n
= 1; Barbour et al., 2011), Northern (n = 3; Fogg et al., 2013) and
Southern Gulf of Mexico (n = 3; Aguilar-Perera \& Quijano-Puerto, 2016),
the Southern Mexican Caribbean (n = 1; Sabido-Itza et al., 2016), Little
Cayman (n = 1; Edwards, Frazer \& Jacoby, 2014), Jamaica (n = 1; Chin,
Aiken \& Buddo, 2016), Bonaire (n = 1; de Leon et al., 2013) and Costa
Rica (n = 1; Sandel et al., 2015) and Fishbase (n = 1; Froese \& Pauly,
2016) were also included. When available, information on sampling
methods, gender differentiation, location, and depth ranges of each
study was retrieved. Whenever gender was not specified, it was assumed
that the results were presented for pooled genders.

@@

All hypothesis testing was performed with an \emph{a priori} confidence
level of \(\alpha = 0.01\) in R version 3.4.0 (R Core Team, 2017). Data
wrangling was done with the tidyverse package (Wickham, 2017). Maps were
created with a mix of functions from the sp (Pebesma \& Bivand, 2005),
rgdal (Bivand, Keitt \& Rowlingson, 2017), tmap (Tennekes, 2017a), and
tmaptools (Tennekes, 2017b) packages. Heteroskedastic-robust standard
errors were calculated with the sandwich (Zeileis, 2004) and lmtest
(Zeileis \& Hothorn, 2002) packages. Models were manipulated with the
broom package (Robinson, 2017). RefManageR (McLean, 2014) was used to
keep track of citations. Raw data and code used in this work is
available at \url{github.com/jcvdav/lionfish_biometry}.

@@

\begin{figure}[htbp]
\centering
\includegraphics{Manuscript_files/figure-latex/histogram figures-1.pdf}
\caption{Kernell density plots for a) Total length (mm) and b) Total
weight (gr) for 109 lionfish sampled in the central Mexican Caribbean.}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\includegraphics{Manuscript_files/figure-latex/scatter plot-1.pdf}
\caption{Length-weight relationship for 109 lionfish sampled in the
central Mexican Caribbean. Points indicate samples, solid line indicates
curve of best fit.}
\end{figure}

@@

From the eight peer-reviewed studies including information on growth
parameters for \emph{P. volitans} and Fishbase (Froese \& Pauly, 2016),
13 parameters were identified (Table III). Two studies (Fogg et al.,
2013; Aguilar-Perera \& Quijano-Puerto, 2016) reported gender-level and
pooled parameters, while the rest presented pooled results. The smallest
coefficient of determination was presented by Chin, Aiken \& Buddo
(2016) with \(R^2 = 0.8715\), while Sabido-Itza et al. (2016) reported
the highest value at \(R^2 = 0.9907\). Reviewed studies presented
information for organisms obtained at depths between 0.5 and 57 m. Two
studies (Aguilar-Perera \& Quijano-Puerto, 2016; Chin, Aiken \& Buddo,
2016) explicitly stated that their organisms were sampled with pole
spears. Five studies (Barbour et al., 2011; Fogg et al., 2013; Edwards,
Frazer \& Jacoby, 2014; Sandel et al., 2015; Sabido-Itza et al., 2016)
mentioned that some of their organisms were obtained with pole spears
(or other type of harpoon). A single study (de Leon et al., 2013) did
not specify how organisms were sampled.

@@

\begin{figure}[htbp]
\centering
\includegraphics{Manuscript_files/figure-latex/review plots-1.pdf}
\caption{Length-weight relationships (n = 14) for eight studies, this
study, and Fishbase. Colors indicate studies from which the parameters
were extracted. Solid lines indicate that the fit was performed for
males and females pooled together. Dotted lines indicate that the
regression was performed on females, and dashed lines indicate it was
performed for males.}
\end{figure}

There were significant differences in expected-to-observed weight ratios
estimated for each pair of parameters (F(13, 1512) = 39.28; p
\textless{} 0.05). From all allometric parameters reviewed, those of
Edwards, Frazer \& Jacoby (2014) provided the lowest weight estimates,
with an expected-to-observed weight ratio of 0.98 \(\pm\) 0.23 (mean
\(\pm\) SD). On the other hand, Barbour et al. (2011) yielded the
highest weight estimates, with a mean (\(\pm\) SD) expected-to-observed
weight ratio of 1.76 \(\pm\) 0.50. Predicted-to-observed weight ratios
and groups identified by Tukey's HSD (\(\alpha = 0.05\)) are presented
in Figure 5.

\begin{figure}[htbp]
\centering
\includegraphics{Manuscript_files/figure-latex/plot comparing predictions-1.pdf}
\caption{. Box and whiskers plot showing the distribution of predicted
to observed biomass ratios for 14 pairs of allometric parameters. Lines
indicate median values, circles indicate mean values, notches represent
95\% confidence intervals arround the median, lower and upper hinges
correspond to the first and third quartiles, whiskers extend to the
largest and lowes values within 1.5 inter-quartile range of the hinge,
small points represent outliers further away than the whiskers. Like
letters indicate values that do not differ significantly (Tukey's HSD; p
\textless{} 0.05).}
\end{figure}

>>>>

%% Submissions for peer-review must enable line-numbering 
%% using the lineno option in the \documentclass command.
%%
%% Preprints and camera-ready submissions do not need 
%% line numbers, and should have this option removed.
%%
%% Please note that the line numbering option requires
%% version 1.1 or newer of the wlpeerj.cls file, and
%% the corresponding author info requires v1.2

\documentclass[fleqn,10pt,lineno]{wlpeerj} % for journal submissions
% \documentclass[fleqn,10pt]{wlpeerj} % for preprint submissions

\title{Allometric growth parameters for invasive lionfish \emph{Pterois
volitans} (Actinopterygii, Scorpaenidae) in the Western Atlantic}

\author[1]{Juan Carlos Villaseñor-Derbez}
\affil[1]{Bren School of Environmental Sciences and Management, University of
California Santa Barbara, Santa Barbara, California, U.S.}
\corrauthor[1]{Juan Carlos Villaseñor-Derbez}{jvillasenor@bren.ucsb.edu}

% \keywords{Keyword1, Keyword2, Keyword3}

\begin{abstract}
Lionfish (\emph{Pterois volitans/miles}) are an invasive species in the
North-Western Atlantic and the Caribbean. In order to better manage the
invasion, inform lionfish removal programs, and estimate biomass
available for harvest, we must be able to accurately estimate their
total biomass, frequently from length observations. This work compares
length-weight relationships of the invasive lionfish through the
invasion range and reports the length--weight relationship for lionfish
in the Central Mexican Caribbean. A review of 13 length--weight
relationships reported in eight peer-reviewed studies and FishBase is
provided. These parameters were used to identify spatial variation in
weight-at-length. For a given length, parameters from the Caribbean
yielded lower weights than those from the Gulf of Mexico and Atlantic,
indicating that weight-at-length is spatially variable. This highlights
the importance of using site-specific parameters to estimate biomass
from length observations. This study also reports a new pair of
length-weight parameters (\(a = 3.2056 imes 10^{-6}; b = 3.235\)) for
organisms sampled in the Central Mexican Caribbean. Findings from this
work can aid managers and decision makers to better select length-weight
parameters when these are not available for their region of interest.
\end{abstract}

\usepackage{amsthm}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}{Lemma}[section]
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{proposition}{Proposition}[section]
\theoremstyle{definition}
\newtheorem{example}{Example}[section]
\theoremstyle{definition}
\newtheorem{exercise}{Exercise}[section]
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\begin{document}

\flushbottom
\maketitle
\thispagestyle{empty}

@@

At least 84\% of the marine eco-regions have reported the presence of an
invasive species \citep{molnar_2008}. These represent a major threat to
local biodiversity and the economic activities that depend on it, like
tourism or fisheries \citep{bax_2003}. Invasive species may also
threaten native species through competition \citep{davis_2003} or
predation. By 2005, the economic cost of invasive species to the United
States was estimated at \$120 billion per year and nearly 42\% of
species that have been included in the Endangered or Threatened species
list have been labeled as such due to presence of invasive species
\citep{pimentel_2005}. This highlights the importance of understanding,
managing, and preventing ecological invasions.

Lionfish (\emph{Pterois volitans/miles} complex) are an invasive species
in the North-Western Atlantic and the Caribbean, likely introduced
through liberation of aquarium-kept organisms \citep{betancurr_2011}.
They are the first marine vertebrates to establish in North Atlantic
\citep{schofield_2009,schofield_2010} and Caribbean coasts
\citep{sabidoitza_2016}. Lionfish have been widely reported in coral
reefs \citep{aguilarperera_2010}, but also in other habitats such as
estuaries \citep{jud_2011}, mangroves \citep{barbour_2010}, areas with
hard-bottoms \citep{muoz_2011}, and mesophotic reefs
\citep{andradibrown_2017}. Due to its threat to local biodiversity, the
speed of their spread, and its difficulty of management, their presence
in these waters has been labeled as a major marine invasion
\citep{hixon_2016}.

A significant amount of research has been done to describe lionfish
feeding ecology in North Carolina \citep{muoz_2011}, the Bahamas
\citep{morris_2009,cote_2013}, Northern Gulf of Mexico
\citep{dahl_2014}, Mexican Caribbean
\citep{valdezmoreno_2012,villaseorderbez_2014}, Belize
\citep{hackerott_2017}, and Costa Rica \citep{sandel_2015}. Their
feeding behavior and high consumption rates can reduce recruitment
\citep{albins_2008} and population sizes \citep{green_2012} of native
reef-fish species, and further the endangerment of critically endangered
reef fish \citep{rocha_2015}. (However, see \citet{hackerott_2017} for a
case where there was no evidence that lionfish affected the density,
richness, or composition of prey fishes). Major efforts have also been
made to understand the possible impacts of the invasion by keeping track
of its range through time \citep{schofield_2009,schofield_2010} and
predicting invasion ranges under climate change scenarios
\citep{grieve_2016}. By combining information from these disciplines,
researchers have been able to predict the trophic impacts of lionfish
\citep{ariasgonzalez_2011}, which can then be translated into
ecosystem-level and economic impacts.

Seeking to reduce lionfish densities, governments and non-profit
organizations have promoted removal programs and incentivized its
consumption \citep{chin_2016}. In some cases, these have shown to
significantly reduce -but not quite eliminate- lionfish abundances at
local scales \citep{sandel_2015,chin_2016,deleon_2013}. The rapid
recovery rates exhibited by lionfish \citep{barbour_2011} and the
persistent populations in mesophotic coral ecosystems
\citep{andradibrown_2017} -which can contribute with recruitment to
shallow-water populations- make of complete eradication through fishing
effort an unlikely solution. However, further incentivizing its
consumption might create a demand big enough to promote and sustain a
stable fishery \citep{chin_2016}, which can reduce local abundances and
control -not eradicate- the invasion while providing alternative
livelihoods.

The feasibility of lionfish removal programs has been extensively
evaluated through field
observations\citep{usseglio_2017,sandel_2015,chin_2016,deleon_2013} and
empirical modeling \citep{barbour_2011,morris_2011,johnston_2015}. The
latter measure changes in biomass or density
\citep{barbour_2011,johnston_2015} in response to increased mortality
(\emph{i.e.} lionfish removal). In this case, biomass represents the sum
of all fish's individual weight. Total Weight (TW) can be estimated from
Total Length (TL) observations using the allometric growth equation
(\(TW = aTL^b\)). Parameters \(a\) and \(b\) for this equation exist for
North Carolina \citep{barbour_2011}, Northern \citep{fogg_2013} and
Southern Gulf of Mexico \citep{aguilarperera_2016}, the Southern Mexican
Caribbean \citep{sabidoitza_2016}, Little Cayman \citep{edwards_2014},
Jamaica \citep{chin_2016}, Bonaire \citep{deleon_2013} and Costa Rica
\citep{sandel_2015}, but remain unavailable for the central Mexican
Caribbean. The weight-at-length of a species can vary across regions as
a response to biotic (\emph{e.g.} local food availability) and abiotic
(\emph{e.g.} water temperature) conditions \citep{johnson_2016}. Thus,
when using biomass-informed models or estimating biomass from length
observations, it is important to use site-specific parameters to obtain
an accurate estimate. This is especially important when research
involves identifying the total biomass available for harvest by fishers
\citep{chin_2016} or the efficacy of lionfish removals
\citep{barbour_2011,morris_2011,johnston_2015}.

@@

The study took place off the coasts of Playa del Carmen, in the Mexican
Caribbean Figure \ref{fig:map}. The region represents the northernmost
section of the Mesoamerican Barrier Reef System \citep{ruizzarate_2004}.
Coral reefs and mangroves are locally important habitats that represent
important sources of income in terms of extractive (\emph{e.g.}
recreational fishing) and non-extractive (\emph{e.g.} SCUBA diving)
activities related to tourism, the main source of income to the local
economy \citep{murray_2007}.

The reef profile has been described by \citet{ariasgonzalez_1998},
indicating that the reef lagoon extends about 500 m from the coast,
until the reef crest is reached. The reef becomes deeper, leading to the
reef front often found at 700 m from the coastline and extends for an
additional 300 m. At approximately 1000 m away from shore and 30 - 40 m
depth, the reef leads to a drop-off. Along a perpendicular profile to
the coast, bands of reef are interrupted by sand patches at 8 - 12 m
deep and 16-18 m deep. Along the coast, these reefs have been reported
to be under significant anthropogenic pressure, likely causing a shift
in structure and function \citep{bozec_2008}.

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-1-1.pdf}
\caption{\label{fig:unnamed-chunk-1}\label{fig:map}Map of the study area.
The black inset on the left (Yucatan Peninsula) indicates the location
where study sites are distributed. On the right, circular markers
indicate sampling sites and the black romboid indicates location of
Puerto Aventuras, Mexico.}
\end{figure}

@@

\begin{table}

\caption{(\#tab:table of locations)Coordinates, minimum, maximum and mean depth (m), and number of samples for each location.}
\centering
\begin{tabular}[t]{llllllr}
\toprule
Location & Lat. & Long. & Min. Depth & Max. Depth & Mean Depth & n\\
\midrule
Canones & 20.477 & -87.233 & 15.0 & 31.2 & 21.6 & 11\\
Castillo & 20.496 & -87.220 & 12.5 & 30.5 & 27.5 & 18\\
Cuevitas & 20.478 & -87.244 & 7.4 & 12.8 & 11.2 & 4\\
Islas & 20.490 & -87.228 & 14.0 & 19.4 & 16.7 & 10\\
Paamul & 20.513 & -87.192 & 9.9 & 22.7 & 15.5 & 31\\
\addlinespace
Paraiso & 20.484 & -87.226 & 9.4 & 38.1 & 17.7 & 16\\
Pared & 20.502 & -87.212 & 12.1 & 21.0 & 16.3 & 12\\
Pedregal & 20.507 & -87.204 & 14.4 & 14.9 & 14.7 & 3\\
Santos & 20.493 & -87.222 & 5.7 & 26.6 & 16.2 & 2\\
Tzimin-Ha & 20.393 & -87.307 & 21.2 & 24.6 & 22.9 & 2\\
Total &  &  & 5.7 & 38.1 & 18.6 & 109\\
\bottomrule
\end{tabular}
\end{table}

@@

\begin{equation}
\label{eq:allometric}
TW = aTL^b
\end{equation}

@@

\begin{equation}
\label{eq:log-alo}
log_{10}(TW) = b\times log_{10}(TL) + log{10}(a)
\end{equation}

@@

\begin{equation}
\label{eq:log-alo-trans}
Y = mX + c
\end{equation}

@@

Other length-weight relationships (n = 13) were extracted from
peer-reviewed literature. Parameters were obtained for North Carolina
{[}n = 1,barbour\_2011{]}, Northern {[}n = 3,fogg\_2013{]} and Southern
Gulf of Mexico {[}n = 3,aguilarperera\_2016{]}, the Southern Mexican
Caribbean {[}n = 1,sabidoitza\_2016{]}, Little Cayman {[}n =
1,edwards\_2014{]}, Jamaica {[}n = 1,chin\_2016{]}, Bonaire {[}n =
1,deleon\_2013{]} and Costa Rica {[}n = 1,sandel\_2015{]} and Fishbase
{[}n = 1,froese\_website\_2016{]} were also included. When available,
information on sampling methods, gender differentiation, location, and
depth ranges of each study was retrieved. Whenever gender was not
specified, it was assumed that the results were presented for pooled
genders.

@@

All hypothesis testing was performed with an \emph{a priori} confidence
level of \(\alpha = 0.01\) in R version 3.4.0 \citep{rcore_2017}. Data
wrangling was done with the tidyverse package \citep{tidyverse_2017}.
Maps were created with a mix of functions from the sp \citep{sp_2017},
rgdal \citep{rgdal_2017}, tmap \citep{tmap_2017}, and tmaptools
\citep{tmaptools_2017} packages. Heteroskedastic-robust standard errors
were calculated with the sandwich \citep{sandwich_2014} and lmtest
\citep{lmtest_2002} packages. Models were manipulated with the broom
package \citep{broom_2017}. Raw data and code used in this work is
available at github.com/jcvdav/lionfish\_biometry.

@@

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/histogram figures-1.pdf}
\caption{(\#fig:histogram figures)Kernell density plots for a) Total
length (mm) and b) Total weight (gr) for 109 lionfish sampled in the
central Mexican Caribbean.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/scatter plot-1.pdf}
\caption{(\#fig:scatter plot)Length-weight relationship for 109 lionfish
sampled in the central Mexican Caribbean. Points indicate samples, solid
line indicates curve of best fit.}
\end{figure}

@@

From the eight peer-reviewed studies including information on growth
parameters for \emph{P. volitans} and Fishbase
\citep{froese_website_2016}, 13 parameters were identified (Table III).
Two studies \citep{aguilarperera_2016,fogg_2013} reported gender-level
and pooled parameters, while the rest presented pooled results. The
smallest coefficient of determination was presented by \citet{chin_2016}
with \(R^2 = 0.8715\), while \citet{sabidoitza_2016} reported the
highest value at \(R^2 = 0.9907\). Reviewed studies presented
information for organisms obtained at depths between 0.5 and 57 m. Two
studies \citep{aguilarperera_2016,chin_2016} explicitly stated that
their organisms were sampled with pole spears. Five studies
\citep{sandel_2015,barbour_2011,fogg_2013,edwards_2014,sabidoitza_2016}
mentioned that some of their organisms were obtained with pole spears
(or other type of harpoon). A single study \citep{deleon_2013} did not
specify how organisms were sampled.

@@

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/review plots-1.pdf}
\caption{(\#fig:review plots)Length-weight relationships (n = 14) for
eight studies, this study, and Fishbase. Colors indicate studies from
which the parameters were extracted. Solid lines indicate that the fit
was performed for males and females pooled together. Dotted lines
indicate that the regression was performed on females, and dashed lines
indicate it was performed for males.}
\end{figure}

There were significant differences in expected-to-observed weight ratios
estimated for each pair of parameters (F(13, 1512) = 39.28; p
\textless{} 0.05). From all allometric parameters reviewed, those of
\citet{edwards_2014} provided the lowest weight estimates, with an
expected-to-observed weight ratio of 0.98 \(\pm\) 0.23 (mean \(\pm\)
SD). On the other hand,barbour\_2011 yielded the highest weight
estimates, with a mean (\(\pm\) SD) expected-to-observed weight ratio of
1.76 \(\pm\) 0.50. Predicted-to-observed weight ratios and groups
identified by Tukey's HSD (\(\alpha = 0.05\)) are presented in Figure 5.

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/plot comparing predictions-1.pdf}
\caption{(\#fig:plot comparing predictions). Box and whiskers plot
showing the distribution of predicted to observed biomass ratios for 14
pairs of allometric parameters. Lines indicate median values, circles
indicate mean values, notches represent 95\% confidence intervals
arround the median, lower and upper hinges correspond to the first and
third quartiles, whiskers extend to the largest and lowes values within
1.5 inter-quartile range of the hinge, small points represent outliers
further away than the whiskers. Like letters indicate values that do not
differ significantly (Tukey's HSD; p \textless{} 0.05).}
\end{figure}
